\documentclass[12pt]{article}

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{cite}
\usepackage{enumerate}
\usepackage{fancyhdr}
\usepackage[headheight=1in,margin=1.25in]{geometry}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{mathtools}
\usepackage{scalerel}
\usepackage{setspace}

\usepackage{tikz}

\newcommand{\N}{\ensuremath{\mathbb{N}}}
\newcommand{\Z}{\ensuremath{\mathbb{Z}}}
\newcommand{\Q}{\ensuremath{\mathbb{Q}}}
\newcommand{\R}{\ensuremath{\mathbb{R}}}
\newcommand{\C}{\ensuremath{\mathbb{C}}}

\newcommand{\e}{\ensuremath{\varepsilon}}
\renewcommand{\d}{\ensuremath{\delta}}

\newcommand{\angleb}[1]{\left\langle#1\right\rangle}
\newcommand{\braceb}[1]{\left\{#1\right\}}
\newcommand{\bracketb}[1]{\left[#1\right]}
\newcommand{\parenb}[1]{\left(#1\right)}
\newcommand{\vertb}[1]{\left\vert#1\right\vert}
\newcommand{\dvertb}[1]{\left\Vert#1\right\Vert}

\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}

\renewcommand{\Re}{\text{Re}}
\renewcommand{\Im}{\text{Im}}

\newcommand{\comp}{\complement}
\newcommand{\sdiff}{\setminus}

\newcommand{\solution}{\textit{Solution: }}
\newcommand{\proof}{\textit{Proof: }}
\newcommand{\partialdone}{\ensuremath{\strut\hfill\blacktriangle}}
\newcommand{\done}{\ensuremath{\strut\hfill\blacksquare}}

\newcommand{\mc}[1]{\ensuremath{\mathcal{#1}}}

\renewcommand{\t}[1]{\text{ #1 }}
\newcommand{\impl}{\ensuremath{\implies}}
\newcommand{\sectionskip}{\vspace{0.15in}}
\newcommand{\tri}{\triangle}
\newcommand{\ovl}[1]{\ensuremath{\overline{#1}}}

\begin{document}

\pagestyle{fancy}
\fancyhead[L]{Samuel}
\fancyhead[C]{Algebraic Number Theory}
\fancyhead[R]{Chapter 1}

\setlength{\parindent}{0in}
\setlength{\parskip}{0.1in}
\setstretch{1}

\section*{Preliminaries}

We define the natural numbers \( \N \) to be the strictly positive integers.
That is, \( \N = \braceb{n \in \Z : n > 0} \).
We additionally define \( \N_0 = \N \cup \braceb{0} \).

Unless otherwise stated, all rings are commutative and contain a 1.

Given a ring \( R \) and two ideals \( I \) and \( J \) of \( R \), define
their sum and product as follows:
\[
	I + J = \braceb{a + b : i \in a \t{and} b \in J},
\]
and
\[
	IJ = \braceb{
		\sum_{k = 1}^n a_kb_k : a_k \in I, b_k \in J, \t{and} n \in \N
	}.
\]
We can generalize the product of ideals to finite collections of ideals.
Let \( I_1, I_2, \dots, I_m \) be ideals of a ring \( R \), then we define
their product as follows:
\[
	I_1I_2 \cdots I_m = \braceb{
		\sum_{k = 1}^n \parenb{\prod_{l = 1}^m a_{k,l}}
		: a_{k,l} \in I_k \t{and} n \in \N
	}
\]
As a special case, the power ideal \( I^m \) is defined for \( m \in \N \) as
follows:
\[
	I^m = \braceb{
		\sum_{k = 1}^n \parenb{\prod_{l = 1}^m a_{k,l}}
		: a_{k,l} \in I \t{and} n \in \N
	}
\]

Given a ring \( R \) and an arbitrary indexing set \( I \), we define
\( R^{(I)} \) as the set of collections \( \braceb{a_i}_{i \in I} \) of
elements of \( R \) such that \( a_i = 0 \) for almost all \( i \in I \).
We can endow this set with an \( R \)-module structure by defining
componentwise addition and scalar multiplication.

We define \( \d_{i,j} \) as follows:
\[
	\d_{i,j} = \begin{cases}
		1 & \t{if} i = j  \\
		0 & \t{otherwise}
	\end{cases}
\]
Additionally, for the \( R \)-module \( R^{(I)} \), we define the elements
\( e_i = \braceb{d_{i,j}}_{j \in I} \) and denote
\( \braceb{e_i}_{i \in I} \) as the canonical basis for \( R^{(I)} \).

Given an \( R \)-module \( M \), it is understand that \( 0 \) serves as the
identity element of the abelian group \( M \).
Additionally, we denote \( L(M,A) \) as the set of linear forms on \( M \).
That is, \( L(M,R) \) is the set of functions \( f : M \to R \) such that
\( f(m + n) = f(m) + f(n) \) and \( f(am) = af(m) \) for all \( m, n \in M \)
and \( a \in R \).

\section*{1.1 Divisibility in Principal Ideal Rings}

\textbf{Proposition 1.1:}
\( \Z \) is a principal ideal domain.

\proof
Let \( I \) be a nonzero ideal in \( \Z \), then we can choose a nonzero
element \( b \in I \).
If \( b < 0 \), then \( b^2 > 0 \) and \( I \) has a positive element, thus
without loss of generality we can assume \( b > 0 \).
By the well ordering principle, we can also assume that \( b \) is minimal
among the positive integers in \( I \).
Now, fix an arbitrary element \( x \in I \).
Since \( \Z \) is a Euclidean domain, we have unique integers \( q \) and
\( r \) such that \( x = qb + r \), and where \( 0 \leq r < b \).
Consequently, we have that \( r = x - qb \in I \) since \( I \) is closed under
addition.
If \( r > 0 \), then we have found a positive integer in \( I \) that is
less than \( b \), a contradiction to our assumption, hence we must have
\( r = 0 \).
This shows that all elements in \( I \) are multiples of \( b \), and thus
\( I = (b) \) and is a principal ideal.
\done

\textbf{Proposition 1.2:}
If \( K \) is a field, then \( K[X] \) is a principal ideal domain.

\proof
Fix a nonzero ideal \( I \) in \( K[X] \).
Let \( d \) be the minimal nonzero degree of any element in \( I \), and let
\( b \) be an element in \( I \) with degree \( d \).
Letting \( x \) be any element in \( I \), we can use the fact that \( K[X] \)
is a Euclidean domain to find unique elements \( q, r \in I \) where
\( x = qb + r \), and where \( 0 \leq \deg(r) < \deg(b) \).
By closure, we have that \( r = x - qb \in I \), which combined with our
assumption on \( \deg(b) \) forces \( \deg(r) = 0 \).
Thus, any element \( x \in I \) is a multiple of \( b \), meaning that
\( I = (b) \) and is thus principal.
\done

\section*{1.2 Diophantine Equations}

\textbf{Theorem 1.3:}
We have that \( x, y, z \in \N \) satisfy the equation \( x^2 + y^2 = z^2 \)
if and only if there exists an integer \( d \) and relatively prime integers
\( u \) and \( v \) such that, after possible rearrangement of \( x \) and
\( y \):
\[
	x = d(u^2 - v^2), \quad
	y = 2duv, \quad\text{and}\quad
	z = d(u^2 + v^2).
\]

\proof
The backwards direction is easy to see, as
\[
	x^2 + y^2
	= [d(u^2 - v^2)]^2 + (2duv)^2
	= d^2(u^4 - 2u^2v^2 + v^4) + d^2(4u^2v^2)
\]
\[
	= d^2(u^4 + 2u^2v^2 + v^4)
	= d^2(u^2 + v^2)^2
	= z^2.
\]
Conversely, let \( x, y, z \in \N \) satisfy the equation.
We can, without loss of generality, assume that they are pairwise relatively
prime, since otherwise we can divide out by their gcd.
\partialdone

\textbf{Theorem 1.4:}
There exist no integers \( x, y, z \in \N \) that satisfy the equation
\[
	x^4 + y^4 = z^2.
\]

\proof
\partialdone

\textbf{Corollary 1.5:}
There exist no integers \( x, y, z \in \N \) that satisfy the equation
\[
	x^4 + y^4 = z^4.
\]

\proof
\partialdone

\section*{1.3 Lemmas on Ideals and Euler's \( \phi \)-function}

\textbf{Proposition 1.6:}
Fix natural numbers \( q \) and \( n \), and denote \( \tilde{q} \) as the
residue class \( q + n\Z \).
We have that the following are equivalent:
\begin{enumerate}[(a)]
	\item \( \gcd(q,n) = 1 \)

	\item \( \tilde{q} \) is a unit in the ring \( \Z/n\Z \)

	\item \( \tilde{q} \) generates the additive group \( \Z/n\Z \)
\end{enumerate}
As an immediate consequence, we have that \( \phi(n) \) is equal to the number
of units in the ring \( \Z/n\Z \), as well as the number of generators of the
additive group \( \Z/n\Z \).

\proof
We will prove that (a) \impl (b) \impl (c) \impl (a).
We also choose \( q \) to be the unique representative of \( \tilde{q} \)
where \( 0 \leq q \leq n - 1 \).

First, assume (a).
By Bezout's lemma, there exist integers \( x \) and \( y \) such that
\( qx + ny = 1 \).
We thus have that \( qx \equiv 1 \pmod{n} \), which is equivalent to
\( \tilde{q} \cdot \tilde{x} = \tilde{1} \).
This proves (a) \impl (b), and so we now assume that \( \tilde{q} \) is a unit
in \( \Z/n\Z \).
Choose \( \tilde{x} \in \Z/n\Z \) such that
\( \tilde{q} \cdot \tilde{x} = \tilde{1} \), then for arbitrary
\( \tilde{a} \in \Z/n\Z \) we have
\( \tilde{a} = \tilde{a} \cdot \tilde{x} \cdot \tilde{q} \).
This is equivalent to \( \tilde{a} = (ax)\tilde{q} \), which shows that
\( \tilde{q} \) generates all elements in \( \Z/n\Z \), and (b) \impl (c).
Finally, assume \( \tilde{q} \) to be a generator for \( \Z/n\Z \), then
there exists an integer \( x \) where \( x\tilde{q} = \tilde{1} \), but this
means that \( xq \equiv 1 \pmod{n} \).
We can choose an integer \( y \) such that \( xq - 1 = yn \), which implies
that \( xq - yn = 1 \).
By Bezout's lemma, this implies that \( \gcd(q,n) = 1 \), thus showing
(c) \impl (a) and completing the proof.
\done

\textbf{Lemma 1.7:}
Let \( R \) be a ring and let \( I \) and \( J \) be ideals of \( R \) such
that \( I + J = R \), then we have that \( I \cap J = IJ \) and
\( R/IJ \cong R/I \times R/J \).

\proof
We have that \( IJ \subseteq I \) since \( I \) absorbs multiplication from
\( J \).
Similarly, \( IJ \subseteq J \), and thus \( IJ \subseteq I \cap J \).
Fix an element \( x \in I \cap J \).
By assumption, \( I + J = R \), and so there exist elements \( a \) and \( b \)
in \( I \) and \( J \), respectively, such that \( a + b = 1 \), hence we
obtain the equality \( x = xa + xb \).
This means that \( x \) is an element of \( IJ \), and thus
\( IJ \subseteq I \cap J \).
A a result, we have \( IJ = I \cap J \).

Next, we consider the ring homomorphism \( \theta : R \to R/I \times R/J \),
which we define as \( \theta(a) = (a + I, a + J) \).
Since \( a + I = I \) and \( a + J = J \) if and only if \( a \in I \cap J \),
we have that \( \ker\theta = I \cap J = IJ \).
This shows that, if \( a + IJ \) and \( b + IJ \) are equivalent elements of
\( R/IJ \), then \( \theta(a) = \theta(b) \).
We can thus define a function \( \phi : R/IJ \to R/I \times R/J \), defined
as \( \phi(a + IJ) = \theta(a) = (a + I, a + J) \).
It is easy to verify that \( \phi \) is a homomorphism.
Additionally, if \( a \in \ker\theta \), then
\( \phi(a + IJ) = \theta(a) = (I,J) \), which shows that
\( \ker\phi = \braceb{0 + IJ} \), and thus \( \phi \) is injective.
Finally, fix \( (y + I, z + J) \in R/I \times R/J \), and again take \( a \)
and \( b \) as elements in \( I \) and \( J \), respectively, such that
\( a + b = 1 \).
Defining the element \( x \in R \) as \( x = az + by \), we can see that
\[
	x \equiv by \equiv (1 - a)y \equiv y - ay \equiv y \pmod{a},
\]
as well as
\[
	x \equiv az \equiv (1 - b)z \equiv z - bz \equiv z \pmod{b}.
\]
This shows that
\[
	\phi(x + IJ) = \theta(x) = (x + I, x + J) = (y + I, z + I),
\]
thus \( \phi \) is surjective, hence an isomorphism, and
\( R/IJ \cong R/I \times R/J \).
\done

\textbf{Lemma 1.8:}
Let \( R \) be a ring and let \( I_1, I_2, \dots, I_n \) be ideals of \( R \)
such that \( I_i + I_j = R \) for all \( 1 \leq i < j \leq n \).
We have that
\( A/I_1I_2 \cdots I_n \cong A/I_1 \times A/I_2 \times \cdots \times A/I_n \).

\proof
The previous lemma proves the case for \( n = 2 \).
We now use induction on \( n \) and assume the case for \( n - 1 \) holds.
Define the ideal \( J = I_2I_3 \cdots I_n \) in \( R \).
Note that for \( 2 \leq k \leq n \), we have \( J \subseteq I_k \) and
\( I_1 + I_k = R \), thus we can find elements \( a_k \in I_1 \) and
\( b_k \in J \) such that \( a_k + b_k = 1 \), and thus
\[
	1 = \prod_{k = 2}^n (a_k + b_k).
\]
Using this we obtain the equality \( 1 = c + a_2a_3 \cdots a_n \), where
\( c \) is the sum of the terms in the product that contain at least one
\( b_k \).
This implies that \( c \in I_1 \), and thus
\( 1 = c + a_2a_3 \cdots a_n \in I_1 + J \).
Because \( I_1 + J \) is an ideal that contains 1, we know that
\( I_1 + J = R \).

By the previous lemma, we have
\[
	A/I_1J \cong A/I_1 \times A/J,
\]
and invoking the induction hypothesis, we can see that
\[
	A/J
	= A/I_2I_3 \cdots I_n
	\cong A/I_2 \times A/I_3 \times \cdots \times A/I_n,
\]
thus we have that
\[
	A/I_1I_2 \cdots I_n
	= A/I_1J
	\cong A/I_1 \times A/J
	\cong A/I_1 \times A/I_2 \times \cdots \times A/I_n,
\]
completing the proof.
\done

\textbf{Proposition 1.9:}
Let \( m \) and \( n \) be relatively prime integers, then we have that
\( \Z/(mn)\Z \cong \Z/m\Z \times \Z/n\Z \).

\proof
By Bezout's lemma, there exist integers \( x \) and \( y \) such that
\[
	xm + yn = 1,
\]
which means that the ideal \( m\Z + n\Z \) contains 1, and thus is equal
to \( \Z \).
Applying \textbf{Lemma 1.7}, we obtain
\[
	\Z/(mn)\Z
	= \Z/(m\Z n\Z)
	\cong \Z/m\Z \times \Z/n\Z.
\]
\done

\textbf{Corollary 1.10:}
If \( m \) and \( n \) are relatively prime, then
\( \phi(mn) = \phi(m)\phi(n) \).

\proof
We previously established that \( \phi(mn) \) is equal to the number of units
in \( \Z/mn\Z \), and thus the number of units in \( \Z/m\Z \times \Z/n\Z \).
Since an element \( (a,b) \in \Z/m\Z \times \Z/n\Z \) is a unit if and only if
\( a \) and \( b \) are units in their respective rings, we have exactly
\( \phi(m)\phi(n) \) choices for units in \( \Z/m\Z \times \Z/n\Z \), thus
proving the equality.
\done

\textbf{Corollary 1.11:}
For fixed \( n \in \N \), let \( p_1^{a_1}p_2^{a_2} \cdots p_m^{a_m} \) be its
prime factorization, then we have that
\[
	\phi(n) = n\prod_{k = 1}^m \parenb{1 - \frac{1}{p_k}}.
\]

\proof
Since powers of distinct primes are always relatively prime, we have that
\( p_i^{a_i}\Z + p_j^{a_j}\Z = \Z \) for all \( 1 \leq i < j \leq m \).
Additionally, \( \Z/n\Z \) has \( \phi(n) \) units.
Furthermore, we have that
\[
	\Z/n\Z
	\cong \Z/p_1^{a_1}\Z p_2^{a_2}\Z \cdots p_m^{a_m}\Z
	\cong \Z/p_1^{a_1}\Z
	\times \Z/p_2^{a_2}\Z
	\times \cdots \times \Z/p_m^{a_m}\Z,
\]
but the number of units in the right hand side is equal to
\( \phi(p_1^{a_1})\phi(p_2^{a_2}) \cdots \phi(p_m^{a_m}) \).
Thus, we have that
\[
	\phi(n)
	= \prod_{k = 1}^m \phi(p_k^{a_k})
	= \prod_{k = 1}^m p_k^{a_k} - p_k^{a_k - 1}
	= \prod_{k = 1}^m p_k^{a_k}\parenb{1 - \frac{1}{p_k}}
	= \prod_{k = 1}^m p_k^{a_k} \prod_{k = 1}^n \parenb{1 - \frac{1}{p_k}}
\]
\[
	= n\prod_{k = 1}^m \parenb{1 - \frac{1}{p_k}}.
\]
\done

\section*{1.4 Preliminaries on Modules}

\textbf{Proposition 1.12:}
Let \( R \) be a ring, \( I \) an indexing set, and \( M \) an \( R \)-module.
Additionally, let \( \braceb{m_i}_{i \in I} \) be a fixed collection of
elements in \( M \), and let \( a = \braceb{a_i}_{i \in I} \) be an element in
\( R^{(I)} \).
We have that the function \( \phi : R^{(I)} \to M \), defined as
\[
	\phi(a) = \sum_{i \in I} a_im_i,
\]
is an \( R \)-module homomorphism.
Furthermore, the following are true:
\begin{enumerate}[(a)]
	\item \( \braceb{m_i}_{i \in I} \) is linearly independent if and only if
	      \( \phi \) is injective

	\item If \( I \) is finite, then \( \braceb{m_i}_{i \in I} \) generates
	      \( M \) if and only if \( \phi \) is surjective
\end{enumerate}

\proof
Let \( a = \braceb{a_i}_{i \in I} \) and \( b = \braceb{b_i}_{i \in I} \) be
elements in \( R^{(I)} \).
We can see that
\[
	\phi(a + b)
	= \sum_{i \in I} (a_i + b_i)m_i
	= \sum_{i \in I} a_im_i + \sum_{i \in I} b_im_i
	= \phi(a) + \phi(b),
\]
and for \( r \in R \), we have that
\[
	\phi(ra)
	= \sum_{i \in I} ra_im_i
	= r\sum_{i \in I} a_im_i
	= r\phi(a).
\]
Note that we can factor the \( r \) out of the sum since only finitely many
terms are nonzero.
Thus, we have shown that \( \phi \) is an \( R \)-module homomorphism.

Next, assume that the \( m_i \) are linearly independent.
Letting \( a, b \in R^{(I)} \) where \( \phi(a) = \phi(b) \), we see that
\[
	\sum_{i \in I} a_im_i = \sum_{i \in I} b_im_i,
\]
and thus
\[
	\sum_{i \in I} (a_i - b_i)m_i = 0.
\]
The linear independence of the \( m_i \) force \( a_i - b_i = 0 \) for all
\( i \in I \), hence \( a_i = b_i \) and \( a = b \), which shows the
injectivity of \( \phi \).
Conversely, assume \( \phi \) is injective.
It is easy to see that \( \phi(0) = \braceb{0}_{i \in I} \), as
\[
	\phi(0) = \sum_{i \in I} 0m_i = 0.
\]
Additionally, if we have \( a \in R^{(I)} \) such that \( \phi(a) = 0 \), then
by the injectivity of \( \phi \) we have that \( a = 0 \), and thus
\( a_i = 0 \) for all \( i \in I \).

Moving on to (b), we assume that \( \braceb{m_i}_{i \in I} \) generates
\( M \).
Further, we assume that \( I \) is finite.
This means that, for all \( m \in M \), we can represent it as
\( m = \sum_{i \in I} a_im_i \), where the \( a_i \) are elements in \( R \).
Defining \( a = \braceb{a_i}_{i \in I} \), we easily see that
\( \phi(a) = m \), showing the surjectivity of \( \phi \).
Finally, assuming \( \phi \) is surjective, we have for any \( m \in M \) an
element \( a \in R^{(I)} \) such that \( \phi(a) = m \).
This means that
\[
	m = \sum_{i \in I} a_im_i,
\]
thus each element in \( M \) can be represented as a linear combination of the
\( m_i \), showing that \( \braceb{m_i}_{i \in I} \) generates \( M \).
\done

\textbf{Lemma 1.13}
If \( T \) is a partially ordered set, then the following are equivalent:
\begin{enumerate}[(a)]
	\item Every nonempty subset of \( T \) contains a maximal element

	\item Every nondecreasing sequence of elements of \( T \) is stationary
\end{enumerate}

\proof
Fix an increasing sequence of elements in \( T \), say
\( \braceb{t_n}_{n \in \N} \), then assuming (a), we can choose \( t_q \) as
the maximal element of this sequence.
Thus, if \( n \geq q \), then \( t_n \geq t_q \), but the maximality of
\( t_q \) forces \( t_n = t_q \), showing that the sequence is stationary and
that (a) \impl (b).

Conversely, assume (b), and let \( S \) be any nonempty subset of \( T \).
To establish a contradiction, suppose \( S \) has no maximal element, then
for all \( x \in S \), we have some \( y \in S \) where \( x < y \).
Thus, by the axiom of choice, we can define a function \( f : S \to S \) where
all \( x \) map to strictly larger values.
Choosing an arbitrary value \( t_0 \in S \), we can define the sequence
\( \braceb{t_n}_{n \in \N} \) where \( t_n = f(t_{n - 1}) \).
By the construction of \( f \) this is a strictly increasing sequence and thus
cannot be stationary, a contradiction to our assumption.
Thus, we must have that \( S \) has a maximal element, proving (b) \impl (a).
\done

\textbf{Theorem 1.14}

Fix a ring \( R \) and an \( R \)-module \( M \), then the following are
equivalent:
\begin{enumerate}[(a)]
	\item Every collection of submodules of \( M \) has a maximal element
	      under the ordering of set containment

	\item Every nondecreasing sequence of submodules of \( M \) is stationary

	\item Every submodule of \( M \) has finite type
\end{enumerate}

\proof
We show that (a) \impl (c) \impl (b) \impl (a).

First, assuming (a), we fix a submodule \( E \) of \( M \).
Denoting \( \Phi \) as the collection of submodules of \( E \) with finite
type, we let \( F \) be the maximal element of \( \Phi \).
We can do this since \( \braceb{0} \) is a submodule \( E \), and thus
\( \Phi \) is nonempty.
Also note that, for all \( x \in E \), we have by the closure of addition and
scalar multiplication that \( F + Rx \) is a submodule of \( E \), and thus
is of finite type, generated by \( x \) and any finite set of generators for
\( F \).
From this it is clear that \( F \subseteq F + Ax \), and by our assumption of
the maximality of \( F \), we have that \( F = F + Ax \), thus \( x \in F \).
This shows that \( E \subseteq F \), and thus \( E = F \), showing that \( E \)
if of finite type, and proving (a) \impl (c).

Now assume (c), and let \( \braceb{M_n}_{n \in \N} \) be a nondecreasing
sequence of submodules of \( M \).
We define the submodule \( E \) of \( M \) as follows:
\[
	E = \bigcup_{n \in \N} M_n.
\]
By (c), \( E \) is of finite type, and so we can choose a finite set of
generating elements for \( E \), say \( x_1, x_2, \dots, x_q \).
Also, by the definition of \( E \), we have for each \( 1 \leq n \leq q \)
that there exists an index \( k \) such that \( x_n \in M_k \).
Choosing \( x_n \) so that \( k \) is maximal, we see that \( x_n \in M_k \)
for all \( n \) since the sequence is nondecreasing.
Since \( M_k \) contains a generating set of \( E \), we have that
\( M_k = E \).
Additionally, if \( n \geq k \), then \( M_k \subseteq M_n \subseteq E \),
showing that \( M_k = M_n \) and that the sequence is stationary, proving
(c) \impl (b).

Finally, we obtain (b) \impl (a) by noting that it is a special case of
\textbf{Lemma 1.13}.
\done

\textbf{Corollary 1.15:}
If \( R \) is a principal ideal domain, then every collection of ideals
of \( R \) contains a maximal element.

\proof
The ideals of \( R \) are precisely the submodules of \( R \) considered as an
\( R \)-module over itself.
Additionally, as \( R \) is a P.I.D., we have that any ideal (submodule) of
\( R \) is generated by a single element and thus is of finite type.
Invoking (c) \impl (a) from \textbf{Theorem 1.14}, we have shown that any
collection of ideals in \( R \) contains a maximal element.
\done

\section*{1.5 Modules Over Principal Ideal Domains}

\textbf{Theorem 1.16}
Let \( R \) be a principal ideal domain, \( M \) a free \( R \)-module of rank
\( n \), and \( N \) a submodule of \( M \).
Then the following are true:
\begin{enumerate}[(a)]
	\item \( N \) is free and has rank \( q \leq n \)

	\item If \( N \) is not the trivial module, then there exists a basis
	      \( b_1, b_2, \dots, b_n \) of \( M \) and elements
	      \( a_1, a_2, \dots, a_q \in R \) such that
	      \( a_1b_1, a_2b_2, \dots, a_qb_q \) is a basis of \( N \), and where
	      \( a_i \) divides \( a_{i + 1} \) for all \( 1 \leq i \leq q - 1 \).
\end{enumerate}

\proof
This theorem is apparent for the trivial module, and so we can assume that
\( N \) is nontrivial.
Fix \( f \in L(M,R) \) and consider the image \( f(N) \).
Since \( R \) can be interpreted as an \( R \)-module over itself, we have that
\( f \) is an \( R \)-module homomorphism, and thus \( f(N) \) is a submodule
of \( R \).
Consequently, \( f(N) \) is an ideal in \( R \), but since \( R \) is a P.I.D.,
we have that \( f(N) = Ra_f \) for some element \( a_f \in R \) that depends on
\( f \).
By \textbf{Corollary 1.15}, we can choose \( f \) such that \( f(N) \) is
maximal among all ideals \( g(N) \), \( g \in L(M,R) \).

We now choose a basis \( x_1, x_2, \dots, x_n \) of \( M \), and define
\( \pi_i : M \to R \) as the projection function on the
\( \text{i}^{\text{th}} \) coordinate.
This function is well defined since the existence of a basis implies the
existence of a bijective module homomorphism from \( M \) to \( R^n \).
Additionally, because \( N \) is nontrivial, it has at least one element
with a nonzero coordinate, and thus the image \( \pi_i(N) \) is a nontrivial
submodule of \( R \) for at least one \( i \).
Note also that \( \pi_i \in L(M,R) \), thus our assumption of the maximality
of \( f(N) \) gives us \( \pi_i(N) \subseteq f(N) \), showing that
\( f(N) \) is a nontrivial submodule (ideal) of \( R \).
Consequently, the element \( a_f \) is nonzero.

Fix an element \( m \in N \) where \( f(m) = a_f \).
This is possible because \( f(N) = Ra_f \) by construction.
Let \( g \) be an arbitrary linear form in \( L(M,R) \) and define
\( d = \gcd(a_f, g(m)) \).
Since \( d \mid a_f \), we have \( Ra_f \subseteq Rd \), and since
\( Ra_f \) is maximal, we have \( Rd \subseteq Ra_f \), proving
that \( a_f \mid d \), and thus \( a_f \mid g(m) \).

A consequence of this is that \( a_f \mid \pi_i(m) \), thus we can choose
\( b_i \in R \) where \( \pi_i(m) = a_fb_i \) for all \( i \).
Defining \( m' = \sum_{i = 1}^n b_ix_i \), we can see that
\[
	a_fm'
	= \sum_{i = 1}^n a_fb_ix_i
	= \sum_{i = 1}^n \pi_i(m)x_i
	= m.
\]
It is also easy to see that \( a_ff(m') = f(a_fm') = f(m) = a_f \), thus
\( a_f[f(m') - 1] = 0 \), which implies that \( f(m') - 1 = 0 \) since
\( a_f \) is nonzero.
We thus have that \( f(m') = 1 \).

We will now prove the following two assertions:
\begin{enumerate}[(1)]
	\item \( M = \ker f + Rm' \)

	\item \( N = (N \cap \ker f) + Rm \)
\end{enumerate}

\begin{enumerate}[(1):]
	\item For all \( x \in M \), we have \( x = (x - f(x)m') + f(x)m' \),
	      and it is easy to see that \( f(x)m' \in Rm' \).
	      Additionally, we have that
	      \( f(x - f(x)m') = f(x) - f(f(x)m') = f(x) - f(x)f(m') = 0 \),
	      since \( f(m') = 1 \).
	      Thus \( x - f(x)m' \in \ker f \), and as a result
	      \( x \in \ker f + Rm' \).
	      This shows that \( M \subseteq \ker f + Rm' \).
	      The containment \( \ker f + Rm' \subseteq M \) is true since
	      \( \ker f \) and \( Rm' \) consist of elements of \( M \), thus
	      proving \( M = \ker f + Rm' \).

	\item Since \( f(N) = Ra_f \), we have for all \( x \in N \) that
	      \( f(x) = ba_f \) for some \( b \in R \).
	      This means that
	      \( x = (x - f(x)m') + f(x)m' = (x - f(x)m') + ba_fm' \), and thus
	      \( x = (x - f(x)m') + bm \).
	      We already showed that \( x - f(x)m' \in \ker f \), and combined with
	      the fact that \( x - f(x)m' = x - ba_fm' = x - bm \in N \), we have
	      \( x - f(x)m' \in N \cap \ker f \).
	      It has thus been shown that \( x \in (N \cap \ker f) + Rm \), and
	      \( M \subseteq (N \cap \ker f) + Rm \).
	      The converse is clear by recalling the reasoning from (1), thus
	      we have shown (2).
\end{enumerate}

We can now prove (a) by induction on the rank \( q \) of \( N \).
If \( q = 0 \), then \( N = \braceb{0} \) and we are finished.
Otherwise, if \( q > 0 \), we have that the rank of \( N \cap \ker f \)
is \( q - 1 \) by (2), thus \( N \cap \ker f \) is a free \( R \)-module by
the induction hypothesis.
Adding \( m \) to a basis of \( N \cap \ker f \) we obtain a basis for
\( N \), thus showing that \( N \) is a free module and proving (a).

Finally, we show (b) by induction on the rank \( n \) of \( M \).
Again, if \( n = 0 \), then \( M = \braceb{0} \) and we are finished.
For \( n > 0 \), we similarly find that the rank of \( \ker f \) is
\( n - 1 \) using (1).
Using the induction hypothesis on \( \ker f \) and its submodule
\( N \cap \ker f \), we obtain a basis \( y_2, y_3, \dots, y_n \) of
\( \ker f \), an integer \( q \leq n \), and elements
\( a_2, a_3, \dots, a_q \), such that \( a_2y_2, a_3y_3, \dots, a_qy_3 \) is
a basis of \( N \cap \ker f \), and where \( a_i \mid a_{i + 1} \) for
\( 2 \leq i \leq q - 1 \).
We define \( a_1 = a_f \) and \( y_1 = m' \).
By (1), we have that \( y_1, y_2, \dots, y_n \) is a basis for \( M \), and by
(2) combined with the fact that \( m = a_fm' \), we have that
\( a_1y_1, a_2y_2, \dots, a_qy_q \) is a basis for \( N \cap \ker f \).
Define the linear form \( g : M \to R \) where \( g(y_1) = g(y_2) = 1 \)
and \( g(y_i) = 0 \) if \( i \geq 3 \).
We have that
\[
	a_1 = a_f = a_fg(y_1) = g(a_fm') = g(m) \in g(N),
\]
and thus \( Ra_f \subseteq g(N) \).
The maximality of \( Ra_f \) forces \( Ra_1 = Ra_f = g(N) \).
Since \( a_2y_2 \in N \cap \ker f \), we have that
\( a_2 = g(a_2y_2) \in g(N) = Ra_1 \),
which means that \( Ra_2 \subseteq Ra_1 \) and \( a_1 \mid a_2 \), thus
proving (b).
\done

\textbf{Corollary 1.17:}
Fix a principal ideal ring \( R \), and let \( M \) be an \( R \)-module of
finite type, then there exists a finite chain of ideals
\( I_1 \supseteq I_2 \supseteq \cdots \supseteq I_n \) in \( R \) where
\[
	M
	\cong R/I_1 \times R/I_2 \times \cdots \times R/I_n.
\]

\proof
Since \( M \) is of finite type, we can choose a finite generating set which
induces a surjective homomorphism \( \phi : R^n \to M \).
Note that by the first isomorphism theorem for modules, we have that
\( M \cong R^n/\ker\phi \).
The previous theorem allows us to choose a basis \( x_1, x_2, \dots, x_n \)
of \( R^n \), an integer \( q \leq n \), and a basis
\( a_1x_1, a_2x_2, \dots, a_qx_q \) of \( \ker\phi \) where
\( a_i \mid a_{i + 1} \) for \( 0 \leq i \leq q - 1 \).
We thus have that \( R^n \cong Rx_1 \times Rx_2 \times \cdots \times Rx_n \).
Defining \( a_p = 0 \) for \( q + 1 \leq p \leq n \), we obtain
\( \ker\phi \cong Ra_1x_1 \times Ra_2x_2 \times \cdots \times Ra_nx_n \),
and thus
\[
	M
	\cong R^n/\ker\phi
	\cong (Rx_1 \times Rx_2 \times \cdots \times Rx_n)
	/ (Ra_1x_1 \times Ra_2x_2 \times \cdots \times Ra_nx_n)
\]
\[
	\cong R/Ra_1 \times R/Ra_2 \times \cdots \times R/Ra_n.
\]
Finally, since \( a_i \mid a_{i + 1} \), we have that
\( Ra_{i + 1} \subseteq Ra_i \), thus completing the proof.
\done

Note that the fundamental theorem of finitely generated abelian groups,
which states that for any finitely generated abelian group \( G \), there are
integers \( k_1, k_2, \dots, k_r \) and \( n \) where \( k_i \mid k_{i + 1} \)
and
\[
	G
	\cong \Z^n
	\times \Z/\Z k_1 \times \Z/\Z k_2 \times \cdots \times \Z/\Z k_r,
\]
is a
special case of this corollary since any abelian group has a \( \Z \)-module
structure.

\textbf{Corollary 1.18:}
Every torsion free module \( M \) of finite type over a principal ideal domain
\( R \) is free.

\proof
Using the previous result we obtain a chain of ideals
\( I_1 \supseteq I_2 \supseteq \cdots \supseteq I_n \) where
\[
	M \cong R/I_1 \times R/I_2 \times \cdots \times R/I_n.
\]
Assume for the sake of establishing a contradiction that \( I_1 \) is a
nontrivial ideal, then we can choose a nonzero element \( a \in I_1 \).
Fixing a nonzero element \( x + I_1 \in R/I_1 \), we note that
\( (x + I_1, 0, \dots, 0) \) is an element in \( M \).
Finally, since \( ax \) is an element of \( I_1 \), we have that
\( ax + I_1 = I_1 \), and thus
\( a(x + I_1, 0, \dots, 0) = (I_1, 0, \dots, 0) = 0 \),
violating the assumption that \( M \) is torsion free.
We thus have that \( I_1 \) is the trivial ideal, which implies that
\( I_i \) is trivial for all \( i \) since each ideal is contained in
\( I_1 \).
This proves that \( M \cong R^n \), thus permiting the basis
\( (1, 1, \dots, 1) \) and proving that \( M \) is free.
\done

\textbf{Corollary 1.19:}
Let \( R \) be a principal ideal domain, then any \( R \)-module \( M \) of
finite type is isomorphic to a finite product of \( R \)-modules
\( M_1 \times M_2 \times \cdots \times M_n \), where \( M_i \) is either equal
to \( R \) or equal to the quotient \( R/Rp^k \), where \( p^k \) is a power of
some prime element in \( R \).

\proof
Using \textbf{Corollary 1.17}, we find a chain of ideals
\( I_1 \supseteq I_2 \supseteq \dots \supseteq I_n \) of \( R \) where
\( M \) is isomorphic to the product of the quotients of \( R \) by these
ideals.
Now, consider \( R/I_i \) for a fixed \( i \).
If \( I_i \) is trivial, \( R/I_i = R \) and we are finished, so we can assume
that \( I_i \) is nontrivial.
We know that \( I_i = Ra \) for some \( a \in R \), and since \( R \) is a
unique factorization domain, we can choose unique prime elements
\( p_1, p_2, \dots, p_r \in R \), positive integers \( k_1, k_2, \dots, k_r \),
and a unit \( u \in R \) where \( a = up_1^{k_1}p_2^{k_2} \cdots p_r^{k_r} \).
Additionally, since \( \gcd(p_i^{k_i}, p_j^{k_j}) = 1 \) if \( i \ne j \), we
have that \( Rp_i^{k_i} + Rp_j^{k_j} = R \).
Finally, we note the fact that
\[
	Ra
	= Rup_1^{k_1}p_2^{k_2} \cdots p_r^{k_r}
	= Rp_1^{k_1}Rp_2^{k_2} \cdots Rp_r^{k_r},
\]
thus we can invoke \textbf{Lemma 1.8} to see the following:
\[
	R/Ra
	\cong R/Rp_1^{k_1} \times R/Rp_2^{k_2} \times \cdots \times R/Rp_r^{k_r}.
\]
This shows that the product of the \( R/I_i \) consists of factors of \( R \)
or \( R/Rp^k \).
\done

\textbf{Corollary 1.20:}
Let \( G \) be a finite abelian group.
There exists an element \( x \in G \) whose order is the least common multiple
of the orders of the other elements in \( G \).

\proof
Since abelian groups have an induced \( \Z \)-module structure, we invoke
\textbf{Corollary 1.17} to obtain a chain of ideals
\( \Z a_1 \supseteq \Z a_2 \supseteq \cdots \supseteq \Z a_n \) where
\[
	G
	\cong \Z/\Z a_1 \times \Z/\Z a_2 \times \cdots \times \Z/\Z a_n.
\]
We thus have that \( a_i \mid a_{i + 1} \) for \( 1 \leq i \leq n - 1 \).
Additionally, we know that \( a_i \ne 0 \) for all \( i \), otherwise \( \Z \)
would be a factor of \( G \), violating the assumption that \( G \) is finite.
Letting \( x \in \Z/\Z a_n \) where \( x \equiv 1 \pmod{a_n} \), we note that
\( (0, 0, \dots, 0, x) \) is an element in \( G \) and has order \( a_n \).
Fixing an arbitrary element \( y = (y_1, y_2, \dots, y_n) \in G \), we can see
that \( a_ny_i = 0 \) for all \( i \) since \( a_i \mid a_n \).
We thus have that \( a_ny = 0 \), which shows that \( a_n \) is a multiple
of the orders of all elements in \( G \).
Finally, since \( a_n \) is itself the order of an element, we know it is
the least common multiple of these orders.
\done

\section*{1.6 Roots of Unity in a Field}

\textbf{Theorem 1.21:}
Let \( F \) be a field.
If \( G \) is a finite subgroup of \( F^* \), then \( G \) consists of roots of
unity and is cyclic.

\proof
By \textbf{Corollary 1.20}, there exists an element \( x \in G \) whose order,
say \( n \), is a multiple of the order of all elements in \( G \).
This means that every element in \( G \) satisfies the polynomial equation
\( X^n = 1 \), and thus every element is a root of unity.
Since the polynomial \( X^n - 1 \) has at most \( n \) roots, we know
that the order of \( G \) is at most \( n \).
We can also see that \( G \) consists of the \( n \) unique elements
\( x, x^2, \dots, x^{n - 1}, x^n = 1 \), thus \( G \) is generated by
\( x \) and is cyclic.
\done

\section*{1.7 Finite Fields}

\textbf{Proposition 1.22:}
If \( F \) is a field with characteristic \( p \ne 0 \), then \( px = 0 \) for
all \( x \in F \), and \( (x + y)^p = x^p + y^p \) for all \( x, y \in F \).

\proof
The binomial theorem gives us
\[
	(x + y)^p
	= \sum_{k = 0}^p \binom{p}{k} x^ky^{p - k}
	= x^p + y^p + \sum_{k = 1}^{p - 1} \binom{p}{k} x^ky^{p - k}.
\]
Since \( \binom{p}{k} = p!/k!(p - k)! \), and since \( p \) does not divide
\( k!(p - k)! \), we have that it is a multiple of \( p \), and thus each
term in the \( \sum_{k = 1}^{p - 1} \binom{p}{k} x^ky^{p - k} \) is a multiple
of \( p \), and thus equal to 0 since \( F \) has characteristic \( p \).
This shows that \( (x + y)^p = x^p + y^p \), completing the proof.
\done

\textbf{Theorem 1.23:}
Let \( F \) be a finite field with cardinality \( q \), then the following are
true:
\begin{enumerate}[(a)]
	\item The characteristic \( p \) of \( F \) is prime, \( F \) is a vector
	      space of finite dimension \( k \), and \( q = p^k \)

	\item The group \( F^* \) is cyclic and has order \( q - 1 \)

	\item \( x^{q - 1} = 1 \) for all \( x \in F^* \), and \( x^q = x \) for
	      all \( x \in F \).
\end{enumerate}

\proof
We know that \( F \) cannot contain \( \Z \), since \( F \) is finite, thus
it must contain \( \Z/p\Z \) for some prime \( p \).
\( F \) also inherits a \( \Z/p\Z \)-module structure from the unique ring
homomorphism \( \phi : \Z \to F \), and since \( \Z/p\Z \) is a field, \( F \)
is a vector space.
Additionally, since \( F \) is finite, it must have finite dimension \( k \),
which means that \( F \cong (\Z/p\Z)^k \).
Since \( (\Z/p\Z)^k \) has \( p^k \) elements, we have shown that
\( q = p^k \), thus proving (a).
(b) follows from \textbf{Theorem 1.21} since \( F^* \) is finite, and (c)
is immediate from (b).
\done

\textbf{Theorem 1.24:}
Fix a finite field \( F \) and let \( f \) be a homogeneous polynomial of
degree \( d \) in \( F[X_1, X_2, \dots, X_n] \).
If \( d < n \), then there exists a point \( (x_1, x_2, \dots, x_n) \in K^n \)
that is not the origin where \( f(x_1, x_2, \dots, x_n) = 0 \).
In other words, every homogeneous polynomial with coefficients in \( F \) has
a nontrivial root in \( F \).

\proof
Denote \( q \) as the cardinality of \( F \) and \( p \) the characteristic,
then \( q \) is a power of \( p \).
Let \( V \) be the set of zeroes of \( f \) in \( K^n \), i.e. the set
\[
	\braceb{(x_1, x_2, \dots, x_n) \in K : f(x_1, x_2, \dots, x_n) = 0}.
\]
For all \( x \in K^n \sdiff V \), we have by the previous theorem that
\( f(x)^{q - 1} = 1 \), otherwise \( x \) is a root and \( f(x)^{q - 1} = 0 \).
Denoting \( g(x) = f(x)^{q - 1} \), we have shown that \( g \) is the
characteristic function \( \chi_{K^n \sdiff V} \).
We can also see that the number of points in \( K^n \sdiff V \), modulo
\( p \), is given by the sum
\begin{equation}
	\sum_{x \in F^n} g(x).
\end{equation}
If this sum is 0, then \( \vertb{F^n \sdiff V} \) is a multiple of \( p \).
Also note that \( \vertb{F^n} \) is a multiple of \( p \), and thus
\( \vertb{V} \) is a multiple of \( p \) since
\( \vertb{V} = \vertb{F^n} - \vertb{F^n \sdiff V} \).
Because \( p \) divides \( \vertb{V} \), and because the origin is clearly a
root of \( f \), we have that \( \vertb{V} \geq p \geq 2 \), and thus
\( V \) must have a nontrivial root for \( f \).

Consequently, it now suffices to show that the sum in (1) is 0.
Since \( g \) is a linear combination of monomials
\( m(X_1, X_2, \dots, X_n) =X_1^{a_1}X_2^{a_2} \cdots X_n^{a_n} \), it also
suffices to show that
\[
	\sum_{x \in F^n} m(x)
	= \sum_{x \in F^n} x_1^{a_1}x_2^{a_2} \cdots x_n^{a_n}
	= \sum_{x_1 \in F} x_1^{a_1}
	\sum_{x_2 \in F} x_1^{a_2}
	\cdots
	\sum_{x_n \in F} x_1^{a_n}
	= 0
\]
for any monomial \( m \in F[X_1, X_2, \dots, X_n] \), which further reduces
to the problem of solving \( \sum_{x \in F} x^k \) where \( k \in \N_0 \).
If \( k = 0 \), then \( x^k = x^0 = 1 \) for all \( x \in F \), and
thus
\[
	\sum_{x \in F} x^k = \sum_{x \in F} 1 = q \equiv 0 \pmod{p},
\]
so we can take \( k > 0 \), in which case we have that \( 0^k = 0 \), thus
we may restrict the sum to \( F^* \).
\( F \) is finite, and so \( F^* \) is a cyclic group of order \( q - 1 \).
Choosing a generator \( y \in F^* \), we have that the powers of \( y \) from
0 to \( q - 2 \) cover the elements in \( F^* \), and so
\[
	\sum_{x \in F^*} x^k
	= \sum_{i = 0}^{q - 2} (y^i)^k,
	= \sum_{i = 0}^{q - 2} y^{ki},
\]
which is a geometric progression.
We now have two cases:
\begin{enumerate}[(a)]
	\item If \( y^k \ne 1 \), then \( k \) is not a multiple of
	      \( \vertb{y} = q - 1 \), and thus
	      \[
		      \sum_{i = 0}^{q - 2} y^{ki}
		      = \frac{1 - y^{k(q - 1)}}{1 - y^k}
		      = \frac{1 - 1}{1 - y^k}
		      = 0.
	      \]

	\item If \( y^k = 1 \), then \( k \) is a multiple of \( q - 1 \),
	      and so
	      \[
		      \sum_{i = 0}^{q - 2} y^{ki}
		      = \sum_{i = 0}^{q - 2} 1
		      = q - 1.
	      \]
\end{enumerate}
Thus, the sum (1) is 0 unless the \( a_i \) in the representation of \( m \)
are all multiples of \( q - 1 \).
In this case, the degree of \( m \) would be at least \( (q - 1)n \).
However, \( \deg g(x) = \deg f(x)^{q - 1} = (q - 1)d < (q - 1)n \), and so
\( g \) fits the case of (a), and thus (1) is 0, completing the proof.
\done

\end{document}
